---
title: "COVID"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


summary(county_df$mortality.rate)
total_death <- sum(county_df$Death) / sum(county_df$Confirmed) * 100
print(total_death)



us_dates <- as.Date(usc$date, format = "%Y-%m-%d")
us_dates <- unique(us_dates)
length(us_dates)
us_dates
us_date <- data.frame(date = us_dates[order(us_dates)])
head(us_date)
#usc_wide <- us_df
usc$date <- as.Date(usc$date, format = "%Y-%m-%d")
head(usc)
us_df <- usc[usc$county == county[1], ]
us_df <- us_df[-(1:nrow(us_df)), ]

n = 1
for (f in fips) {
  us_date <- data.frame(date = us_dates[order(us_dates)])
  tmp <- as.data.frame(subset(usc, usc$fips == f))
  tmp1 <- merge(us_date, tmp, by = 'date', all.x = TRUE, sort = FALSE)
  tmp1$fips[is.na(tmp1$fips)] <- f
  tmp1$county[is.na(tmp1$county)] <- tmp$county[1]
  tmp1$state[is.na(tmp1$state)] <- tmp$state[1]
  us_df <- rbind(us_df, tmp1)
  print(n)
  n=n+1
  rm(us_date)
}

dim(us_df)
head(us_df)
us_df$cases[is.na(us_df$cases)] <- 0
us_df$deaths[is.na(us_df$deaths)] <- 0
dim(us_df)
length(which(is.na(us_df$fips)))


head(county_df)
length(which(county_df$Confirmed > 10000))
usc_top10000 <- which(county_df$Confirmed > 10000)
county_top10000 <- county_df$Fips[usc_top10000]
county_top10000

us_df_f <- subset(us_df, us_df$fips %in% county_top10000)
dim(us_df_f)
head(us_df_f)

us_df_f$cumulative.cases <- NA
us_df_f$cumulative.death <- NA
for (c in unique(us_df_f$county)) {
  us_df_f$cumulative.cases[us_df_f$county == c] <- cumsum(us_df_f$cases[us_df_f$county == c])
  us_df_f$cumulative.death[us_df_f$county == c] <- cumsum(us_df_f$deaths[us_df_f$county == c])
}

head(us_df_f)
us_df_f$seq <- NA
for (c in unique(us_df_f$county)) {
  s <- which(us_df_f$cumulative.cases[us_df_f$county == c] >= 100)[1]
  us_df_f$seq[us_df_f$county == c] <- s
}
head(us_df_f)
us_df_f <- us_df_f[order(us_df_f$seq), ]
head(us_df_f)
dim(us_df_f)
us_df_f <- us_df_f[, -9]
us_df_f$mortality.rate <- us_df_f$cumulative.death / us_df_f$cumulative.cases * 100
head(us_df_f)
save(us_df_f, file = 'US_county_over_10000_covid19.RData')

## HEATMAP ##

load('./US/US_county_over_10000_covid19.RData')

mypal <-colorRampPalette(c("white", "navy","firebrick4"))(299)
head(us_df_f)
tail(us_df_f)
county_code <- unique(us_df_f$fips)

n = 1
for (c in unique(us_df_f$fips)) {
  tmp_us <- us_df_f[us_df_f$fips == c, ]
  #tmp_us$mortality.rate[tmp_cov$Cumulative.Confirmed < 100] <- 0
  print(dim(tmp_us))
  hm_us <- tmp_us[, c(6,5)]
  hm_us[is.na(hm_us)] <- 0
  mr <- tmp_us$mortality.rate
  
  #ha = HeatmapAnnotation(MR = anno_barplot(mr, ylim = c(0,5),bar_width = 1, gp = gpar(fill = 'hotpink', col = 'hotpink')), annotation_name_side = "left")
  if (n == 1) {
    hm_t <- Heatmap(t(hm_us), cluster_columns = FALSE, cluster_rows = FALSE, col = mypal, show_row_names = FALSE, show_column_names = FALSE, name = 'Number', left_annotation = rowAnnotation(mr = anno_block(gp = gpar(fill = 'gray87', col = 'white'), labels = county_code[n], labels_gp = gpar(col = "black", fontsize = 6)))) #, top_annotation = ha
    assign(paste('hm_t', n, sep = ''), hm_t)
  }
  else {
    hm_t <- Heatmap(t(hm_us), cluster_columns = FALSE, cluster_rows = FALSE, col = mypal, show_row_names = FALSE, show_column_names = FALSE, show_heatmap_legend = FALSE, left_annotation = rowAnnotation(mr = anno_block(gp = gpar(fill = 'gray87', col = 'white'), labels = county_code[n], labels_gp = gpar(col = "black", fontsize = 6)))) #, top_annotation = ha
    assign(paste('hm_t', n, sep = ''), hm_t)
  }
  n = n+1
}


ht_list <- hm_t1 %v% hm_t2 %v% hm_t3 %v% hm_t4 %v% hm_t5 %v% hm_t6 %v% hm_t7 %v% hm_t8 %v% hm_t9 %v% hm_t10 %v% hm_t11 %v%  hm_t12 %v%  hm_t13 %v%  hm_t14 %v%  hm_t15 %v%  hm_t16 %v%  hm_t17 %v% hm_t18 %v% hm_t19 %v% hm_t20 %v% hm_t21 %v% hm_t22 %v% hm_t23 %v% hm_t24 %v% hm_t25 %v% hm_t26 %v% hm_t27 %v% hm_t28 %v% hm_t29 %v% hm_t30 %v% hm_t31 %v% hm_t32 %v% hm_t33 %v% hm_t34 %v% hm_t35 %v% hm_t36 %v% hm_t37 %v% hm_t38 %v% hm_t39 %v% hm_t40 %v% hm_t41 %v% hm_t42 %v% hm_t43 %v% hm_t44 %v% hm_t45 %v% hm_t46 %v% hm_t47 %v% hm_t48 %v% hm_t49 %v% hm_t50 %v% hm_t51 %v% hm_t52 %v% hm_t53

#draw(ht_list,  ht_gap = unit(0, "cm"))

png('US_HeatMap.png', width = 12, height = 16, units = 'in', res = 300)
draw(ht_list,  ht_gap = unit(0, "cm"))
dev.off()

png('US_HeatMap_1.png', width = 12, height = 16, units = 'in', res = 300)
draw(ht_list,  ht_gap = unit(0, "cm"))
dev.off()

```{r Read county data}
require('ggplot2')
require('nlme')
library("rjson")

usc <- read.csv(url("https://github.com/nytimes/covid-19-data/raw/master/us-counties.csv"))
us_density <- read.csv(url("https://github.com/camillol/cs424p3/raw/master/data/Population-Density%20By%20County.csv"))
nextstrain <- read.delim("nextstrain_patient_data.tsv",sep='\t')
nextstrain = nextstrain[nextstrain$country == 'USA',]
nextstrain$gender = nextstrain$sex
nextstrain$gender = NA
nextstrain[(nextstrain$sex=="Male" & !is.na(nextstrain$sex)),"gender"] = 1
nextstrain[(nextstrain$sex=="Female" & !is.na(nextstrain$sex)),"gender"] = 0



#raw_data <- fromJSON(file="ncov_global.json") # I had to save it first. Cannot read directly
#virus_data <- as.data.frame(raw_data$meta)

# fix New York City county, does not have fips
usc[usc$county == "New York City","fips"] = 36061

length(unique(usc$county))
length(unique(usc$fips))

fips <- unique(usc$fips)
fips[c(1:5)]
fips <- na.omit(fips)

head(usc)

fips_l <- list()
county_l <- list()
state_l <- list()
cases_l <- list()
death_l <- list()
density_l <- list() #SVC
doubling_l <- list() #SVC
divergence_l <- list() #SVC
entropy_l <- list()

for (f in fips) {
  tmp <- as.data.frame(subset(usc, usc$fips == f))
  c <- tmp$county[1]
  st <- tmp$state[1]
  s <- sum(tmp$cases)
  d <- sum(tmp$deaths)
  y_death = tmp[tmp$deaths>4,'deaths']
  if (length(y_death)>9) {
    y_death = y_death[1:8]
    x_days = 1:8
    fits <- tryCatch(nls(y_death ~ c*exp(x_days/t), start =c(t=2,c=6)),error = function(e) NA)
    if (is.na(fits)) {doubling = NA
    } else {doubling = as.double(coef(fits)[1])*log(2)}
  } else {
    doubling = NA
  }
  den <- us_density[us_density[,5]==f,]$Density.per.square.mile.of.land.area #SVC
  if (length(den)==0) { den=NA} #SVC
  county_l[[paste(f)]] <- c
  state_l[[paste0(f)]] <- st
  fips_l[[paste0(f)]] <- f
  cases_l[[paste0(f)]] <- s
  death_l[[paste0(f)]] <- d
  density_l[[paste0(f)]] <- den #SVC
  doubling_l[[paste0(f)]] <- doubling #SVC
}

county_df <- data.frame(County = unlist(county_l),
                        State = unlist(state_l),
                        Fips = unlist(fips_l),
                        Confirmed = unlist(cases_l),
                        Death = unlist(death_l),
                        Density = unlist(density_l),
                        Doubling = unlist(doubling_l)) #SVC

county_df <- county_df[order(county_df$Confirmed, decreasing = TRUE),]
county_df$mortality.rate <- county_df$Death / county_df$Confirmed * 100
head(county_df)
state_avg = aggregate(county_df[,c(4,5,6,7,8)],list(county_df$State),mean,na.rm=TRUE)
tmp = aggregate(nextstrain[,c(9,10,13)],list(nextstrain$division),mean,na.rm=TRUE)
state_avg = merge(state_avg,tmp,by="Group.1")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=TRUE}
p <- ggplot(county_df[county_df$Death>10,], aes(x=Density, y=mortality.rate, color=State)) +
  geom_point() +
  scale_x_continuous(trans='log10')+
  scale_y_continuous(trans="log10") +
  xlab('Density (people/square mile)') +
  ylab('Mortality rate')
print(p)

p <- ggplot(county_df[county_df$Death>10,], aes(x=Density, y=Confirmed, color=State)) +
  geom_point() +
  scale_x_continuous(trans='log10')+
  scale_y_continuous(trans="log10") +
  xlab('Density (people/square mile)') +
  ylab('Confirmed Cases')
print(p)

p <- ggplot(county_df[county_df$Death>10,], aes(x=Density, y=Death, color=State)) +
  geom_point() +
  scale_x_continuous(trans='log10')+
  scale_y_continuous(trans="log10") +
  xlab('Density (people/square mile)') +
  ylab('Total Death')
print(p)

p <- ggplot(county_df[county_df$Death>10,], aes(x=Density, y=Doubling, color=State)) +
  geom_point() +
#  geom_boxplot()
  #scale_y_continuous(trans="log10") +
  xlab('Density (people/square mile)') +
  ylab('Doubling (days)')
print(p)

p <- ggplot(county_df[county_df$Death>10,], aes(x=mortality.rate, y=Doubling, color=State)) +
  geom_point() +
#  geom_boxplot()
  #scale_y_continuous(trans="log10") +
  xlab('Mortality Rate') +
  ylab('Doubling (days)') + 
  xlim(0,10)
print(p)

p <- ggplot(state_avg, aes(x=div, y=mortality.rate, color=Group.1)) +
  geom_point() +
#  geom_boxplot()
  #scale_y_continuous(trans="log10") +
  ylab('Mortality Rate') +
  xlab('Divergence') 
print(p)

p <- ggplot(state_avg, aes(x=div, y=Doubling, color=Group.1)) +
  geom_point() +
#  geom_boxplot()
  #scale_y_continuous(trans="log10") +
  ylab('Doubling') +
  xlab('Divergence') 
print(p)

# identify abnormal counties
abnormal_county = county_df$Density<1200 & county_df$Doubling<3.5
ab_county = county_df[abnormal_county,]
ab_county = ab_county[!is.na(ab_county$Death),]
ab_county
abnormal_county = county_df$Density>9000 #& county_df$Doubling>10
ab_county = county_df[abnormal_county,]
ab_county = ab_county[!is.na(ab_county$Death),]
ab_county

```
####---- US Flu data / State ----#####
```{r}
days <-c(1:13) * 7
weekdays <- c(day1, day1 + days)
weekdays
class(usc$date)
head(usc)
usc$DATE <- as.Date(usc$date, format = '%Y-%m-%d')
uscDates <- unique(usc$DATE)
head(usc)

for (i in 1:length(uscDates)) {
  test <- TRUE
  dt <- as.Date(uscDates[i], format = '%Y-%m-%d')
  idx = 1
  while (test == TRUE) {
    if (dt < weekdays[idx]) {
      test = FALSE
      usc$DATE[usc$DATE == dt] <- weekdays[idx-1]
    } else if (idx >= 14) {
      usc$DATE[usc$DATE == dt] <- weekdays[idx]
      test = FALSE
    }
    idx = idx + 1
  }
}
table(usc$DATE)
head(usc)

df_covid_week <- data.frame(State = character(),
                            Cases = integer(),
                            Deaths = integer(),
                            Week = as.Date(character()))

states <- unique(usc$state)
states <- na.omit(states)
for (s in states) {
  tmp1 <- as.data.frame(usc[usc$state == s, ])
  df_tmp <- data.frame(State = character(),
                       Cases = numeric(),
                       Deaths = numeric(),
                       Week = as.Date(character()))
  for (w in 1:length(weekdays)) {
    wd <- as.Date(weekdays[w], format = '%Y-%m-%d')
    tmp2 <- as.data.frame(tmp1[tmp1$DATE == wd, ])
    cases <- max(tmp2$cases)
    death <- max(tmp2$deaths)
    tmp_df <- data.frame(State = tmp2$state[1],
                         Cases = cases,
                         Deaths = death,
                         Week = wd)
    df_tmp <- rbind(df_tmp, tmp_df)
  }
  df_covid_week <- rbind(df_covid_week, df_tmp)
}

head(df_covid_week)

plot(df_covid_week$Cases ~ df_covid_week$Week, pch=19, cex=.5)

nys <- subset(df_covid_week, df_covid_week$State == 'New York')
nys
table(usFlu$SUB.AREA)

usFlu_19_20 <- usFlu[usFlu$SEASON == '2019-20', ]
usFlu_19_20 <- usFlu_19_20[usFlu_19_20$WEEK < 15,]
usFlu_19_20 <- usFlu_19_20[order(usFlu_19_20$SUB.AREA, usFlu_19_20$WEEK), ]

head(usFlu_19_20)
day1 <- as.Date('2020-01-01')
table(usFlu_19_20$WEEK)
days <-c(1:12) * 7
weekdays <- c(day1, day1 + days)
weekdays
usFlu_19_20$weekDay <- rep(weekdays, 52)
head(usFlu_19_20)
save(usFlu_19_20, file = 'US_Flu_report_2019-20.RData')

## Group COVID date into weeks
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
